---
title: "Intro to readxl"
author: "Neil De Luna Olipani"
date: "December 7,2019"
output: html_document
---

# Learning `readxL`

We are working with data in readxl and it's going to be gangster!

We're going to use ppackages:

-tidyverse
-readxl
-writexl
-here

```{r packages}
library(tidyverse) 
library(readxl)
library(writexl)
library(here)




# install.packages("package_name") <!--comment hidden, this comment is hidden but not #install-->

###Read in fish counts
```
```{r}
##### Read in the fish_counts_curated.csv

fish.counts<- read.csv(here("r-workshop-data","fish_counts_curated.csv")) 
```
###Read in data from a single Excel Worksheet 


Now, to read in an excel spreadsheet, we'll use the funcion 'read_excel()', 

substrate_cover_curated.xlsx

```{r}
substrate_cover <- read_excel(here("r-workshop-data","substrate_cover_curated.xlsx"),
                                skip=4,
                                na = "-9999")
                            
                            
```

### Use read_excel() to read in a range of values

```{r}
carp_cover_2000 <- read_excel(here("r-workshop-data","substrate_cover_curated.xlsx"),
                              range = "D12:I15",
                              col_names = c("site",
                                            "transect",
                                            "quad",
                                            "side",
                                            "type",
                                            "coverage"))

```
### How can we read data from an Excel workbook with Multiple work sheets ?




