---
title: "intro-to-ggplot2"
author: "Neil Olipani"
date: "12/8/2019"
output: html_document
---

##Attach Packages 

```{r}
library(tidyverse)
library(readxl)
library(here)
```

## Read in the data

```{r}
ca_np <-read_csv(here("r-workshop-data","ca_np.csv"))

ci_np <- read_xlsx(here("r-workshop-data","ci_np.xlsx"))
```

## Make a graph 

1. Tell R we're using 'ggplot'
2. Tell R whata data we're using, and what variables we want to plot 
3. What type of graph we're creating (geom)

- ```{r, eval=Fasle}
- ggplot(data = This , aes(x= This, y = This))+geom_This()
-aes means aesthetically
- ```


Now let's create a graph of CI visitation:

```{r}
ggplot(data = ci_np, aes (x= year, y = visitors))+
    geom_line()
```
```{r}
ggplot(data = ci_np, aes (x= year, y = visitors))+
    geom_area()
```


## Intro to customizing 'ggplot' graphs

- color= (update point or line color)
- fill = ( update the fill charactersitic of a element)
- linetype = 
- pch = (value)
- size = 
- alpha = 



## geom_line makes the visuals different, aes changes the variable 
```{r}
ggplot(data = ci_np, aes (x= year, y = visitors))+
    geom_line(color = "cyan4", linetype = "dashed")
```
```{r}
ggplot(data = ci_np, aes (x= year, y = visitors))+
  geom_point(color = "cyan4",
             pch = 11,
             size = 4,
             alpa = 0.5)
  
```
I have awesome text here, and below is my amazing finalized graph: 
```{r, echo = FALSE}
ggplot(data = ci_np, aes (x= year, y = visitors))+
  geom_point(color = "cyan4",
             pch = 11,
             size = 4,
             alpa = 0.5)
  
```


## Customize elements based o variables values or outcomes, bigger the population the bigger the point, order of the characteristics


```{r}
ggplot(data = ci_np, aes (x= year, y = visitors))+
  geom_point(aes(size = visitors,
                color = visitors),
                pch = 11,
                color = "cyan4",
                alpha = 0.5)
  
```
Activity: Create a colum plot of CINP visitation over time, where the FILL Color changes based on the number of visitors, and the outline color (color=) is magenta
Reminder AES is a variable value, and outside it is a constant. 

```{r}
ggplot(data = ci_np, aes (x= year, y = visitors))+
  geom_col(aes(fill = visitors),
              color = "magenta")
  
```

```{r}
ggplot(data = ci_np, aes (x= year, y = visitors))+
  geom_line(size = 3, color = "red") +
  theme_dark()
  
```

## Add/update lavels & titles 

Using 'labs'



```{r}
ggplot(data = ci_np, aes (x= year, y = visitors))+
  geom_point(size = 1, 
            color = "nay") +
    theme_minimal()+
  labs (x = "C.R.E.A.M",
        y = "annual visitors",
        title = "Wu Tang",
        subtilte = "American Saga")
  
```

## Combine compatible geoms 


Example: Add a point plot on top of a line plot: 

```{r}
ggplot(data = ci_np, aes (x= year, y = visitors))+
  geom_line( fill= " magenta")+
  geom_point (color = "magenta", 
              aes(size = year))
  
```
```{r}
ggplot(data = ci_np, aes (x= year, y = visitors))+
  geom_line(fill = " magenta")+
  geom_col (color = "magenta", 
              aes(size = year))
  
```
## Creat a line graph with multiple series 

```{r}
ggplot(data = ca_np, aes (x= year, y = visitors, color = park_name))+
  geom_line()

```



## Faceting ggplot graphs 

-Each sereis (group) gets its own plot panel
```{r}
ggplot(data = ca_np, aes (x= year, y = visitors, color = park_name))+
  geom_line(show.legend = FALSE) +
  theme_minimal()+
  facet_wrap (~ park_name)

```


## Use ggsave to export a plot 

```{r}
ggsave ("np_graph.png",dpi = 180, width = 8, height = 7)

```

##sync back to Github remote