---
title: "filter/joins (VLOOKUP in Excel)"
author: "Neil Olipani"
date: "12/8/2019"
output: html_document
---

## Attach Packages 
## Use janitor ::clean_names() to automatically update all of your column names to lowercase_snake_case (this important)
## Why is my pipe operator key board short cut not working? 

```{r, include = FALSE}
library(tidyverse)
library(readxl)
library(here)
```

Read in 3 files 

-invert_counts_curated.xslx
-fish_counts_curated.csv
-kelp_counts_cutated.xlsx

```{r}
invert_counts <- read_xlsx(here("r-workshop-data","invert_counts_curated.xlsx"))
  
  
fish_counts <- read_csv(here("r-workshop-data","fish_counts_curated.csv"))
  
  
kelp_counts_abur <-  read_xlsx(here("r-workshop-data","kelp_counts_curated.xlsx"))
```

- select () function: keep, exclude or rearrange colums 
-filter() function: keeps or excludes ROWS (observations) based on whether or not they match variable conditions
-Chapter 9 


## Filter to match rows based on a character string 

Only keep observations from fish_counts where the common_name variable matches "garibaldi"
## Use janitor ::clean_names() to automatically update all of your column names to lowercase_snake_case (this important)
## == means exact match 
```{r}
fish_garibaldi <- fish_counts %>%
  filter(common_name =="garibaldi")
                
```

fish count over 50
```{r}
fish_over50 <- fish_counts %>%
  filter(total_count >= 50)
```


```{r}
mohawk_fish <- fish_counts %>%
  filter(site=="mohk")
```


"or statement"
```{r}
fish_3sp <- fish_counts %>%
    filter(common_name %in% c("garibaldi","blacksmith","black surfperch"))
```

## Filter to retun rows that match conditions for multiple variable (AND statements)

I want to create a subset called urchin_abur_mohk that only keeps observations where:

-site is "abur" or "mohk" AND 
-species is "purple urchin"


Some ways to do this:

-separate "AND" conditions within a single filter line by a coma


```{r}
urchin_abur_mohk <- invert_counts%>%
  filter(site %in% c("abur","mohk"), common_name == "purple urchin")


##Example Pipe Order 
##pipe_test<-fish_counts %>% 
  ##select(common_name, total_count) %>% 
  ##filter(year ==2016)
```


Activity: create a subset from fish_counts, called low_gb_rw that only contains:

-Observations for garibaldi and rock wrasse
-where the total _count is less than 10 


```{r}
low_gb_rw <- fish_counts %>% 
  filter(common_name %in% c("garibaldi","rock wrasse"),
         total_count < 10)
        
```


##Filter rows that Do Not Match Conditions 

USE != (indicates "does NOT match")

Try to create a subset from fish_counts that does NOT contain observations for Mohawk Reef ("mohk)

```{r}
fish_no_mohk <- fish_counts %>% 
        filter(site != "mohk")
```

An example as part of a piped sequence 

```{r}
senorita_allyear <- fish_counts %>% 
    filter(common_name == "senorita") %>%
    group_by(site) %>%
    summarize(count_all_years = sum(total_count))
```


## A couple ways pf joining data frames 

- 'full_join()': keeps everything (safe join)
- 'left_join()': : If we merging A and B, will only retain observations from B that have matches A 
   

```{r}   
fish_2016_2017 <- fish_counts %>% 
  filter(year==2016 | year==2017)
```




