---
title: "intro-to-ggplot2"
author: "Neil Olipani"
date: "12/8/2019"
output: html_document
---

##Attach Packages 

```{r}
library(tidyverse)
library(readxl)
library(here)
```

## Read in the data

```{r}
ca_np <-read_csv(here("r-workshop-data","ca_np.csv"))

ci_np <- read_xlsx(here("r-workshop-data","ci_np.xlsx"))
```

## Make a graph 

1. Tell R we're using 'ggplot'
2. Tell R whata data we're using, and what variables we want to plot 
3. What type of graph we're creating (geom)

- ```{r, eval=Fasle}
- ggplot(data = This , aes(x= This, y = This))+geom_This()
-aes means aesthetically
- ```


Now let's create a graph of CI visitation:

```{r}
ggplot(data = ci_np, aes (x= year, y = visitors))+
    geom_line()
```
```{r}
ggplot(data = ci_np, aes (x= year, y = visitors))+
    geom_area()
```


## Intro to customizing 'ggplot' graphs

- color= (update point or line color)
- fill = ( update the fill charactersitic of a element)
- linetype = 
- pch = (value)
- size = 
- alpha = 



## geom_line makes the visuals different, aes changes the variable 
```{r}
ggplot(data = ci_np, aes (x= year, y = visitors))+
    geom_line(color = "cyan4", linetype = "dashed")
```
```{r}
ggplot(data = ci_np, aes (x= year, y = visitors))+
  geom_point(color = "cyan4",
             pch = 11,
             size = 4,
             alpa = 0.5)
  
```
I have awesome text here, and below is my amazing finalized graph: 
```{r, echo = False}
ggplot(data = ci_np, aes (x= year, y = visitors))+
  geom_point(color = "cyan4",
             pch = 11,
             size = 4,
             alpa = 0.5)
  
```

