---
title: "Pivot tables"
author: "Neil Olipani"
date: "12/8/2019"
output: html_document
---

## Attach Libraries 

```{r}
library(tidyverse)
library (readxl)
library(here)
library(skimr) #install.packages("skimr")

```

## Read in our daa 

Our data is about lobsters 

#changing lobster to lobster2 

```{r}
lobsters <- read_xlsx(here("r-workshop-data","lobsters2.xlsx"))
```


Explore our data

```{r}

#head shows the first 6 lines of data
head(lobsters)

#summary () to give a summary of each column 
summary(lobsters)

## skimr: skim() gives additional summary statistics and is pipable ( %>% )
  skim(lobsters)
  
##ggplot of lobsters 
  ggplot(data =lobsters, aes(x = year))+ 
    geom_bar()+
    facet_wrap(~site)

```

## Our Task:

Hoe has the average size of lobsters has changed over time across times?

Pivot Tables in R are group_by %>% summarize 

```{r}
lobsters %>% # %>% means "and then" lobsters and then by group by year and then etc.
  group_by(year) %>% 
  summarize(count_by_year=n() )
```

What if we didn't group_by first?

```{r}
lobsters %>% 
  summarize(count_by_something = n())

##assign our summary to a new variable 

lobsters_sum <- lobsters %>% 
  summarize(count_by_something = n())

  
```

## group_by multiple variables
And summarize by multiple variables 
We used mean() to calculate the mean size of lobsters by each site and year, and used na.rm = TRUE to calcilate mean despite having missing values in our raw data


```{r}
siteyear_summary <- lobsters %>% 
  group_by(site, year) %>% 
  summarize(count_by_siteyear = n(), 
            mean_size_m = mean(size_mm, na.rm = TRUE),
            sd_size_mm = sd(size_mm), na.rm=TRUE)

siteyear_summary
```
## Activity 

 Calculate the median size_mm(?median)
 create and ggsave() a plot 
 
```{r}
siteyear_summary <- lobsters %>% 
  group_by(site, year) %>% 
  summarize(count_by_siteyear = n(), 
            mean_size_m = mean(size_mm, na.rm = TRUE),
            sd_size_mm = sd(size_mm, na.rm=TRUE),
            median_size_mm = median(size_mm, na.rm = TRUE))

siteyear_summary
```
 

##ggplot: geom_line 

```{r}
ggplot (data = siteyear_summary, aes(x = year, y = median_size_mm, 
    color = site))+geom_line()
  
  ggsave(here("figures","lobsters-line.png"))
```





#mutate () add a new column 
it will make a new variable 

-add a new column that is the size in meres 
- add a colum that is a repeated value 

```{r}
head (lobsters)

lobsters %>% 
    mutate(size_m = size_mm/1000,
           millenia = 2000,
           observer = "Allison Horst")
```

## select() function to subset 

let's selet 
-date 
- site 
-size_ml

```{r}
lobsters_selected <- lobsters %>%
  mutate(size_m = size_mm/1000) %>% 
  select(date, site, size_m)
```






